A Sound Deep Embedding of Arbitrary Normal Modal Logics in Coq

### 2022-09-05 paper/paper.tex
Fix a few more typos

Formally, we can only interpret a formula with respect to a \emph{model}, which is
defined as follows.
A \emph{frame} $\mathcal{F} = \langle W, R \rangle$ consists of a pair formed by
a non-empty set of worlds, $W$, and a binary accessibility relation between
worlds, $R \subseteq W \times W$. A pair $w_1Rw_2$ denotes that world $w_2$ is
accessible from $w_1$. A model $\mathcal{M} = \langle \mathcal{F}, \mathcal{V} \rangle$
is formed by a frame $\mathcal{F}$ and a total labeling function
$\mathcal{V} : \mathcal{P}\to 2^W$, which assigns
a truth value to a propositional variable at a given world, i.e. when $w \in \mathcal{V}(A)$
for some $A \in\mathcal{P}$, we say that $A$ holds in $w$. The satisfiability of a formula
in a given world is defined by inductive relation $\mathcal{M}\,;\,w\models \varphi$ as
follows (notation $\mathcal{M}\,;\,w\not\models \varphi$ denotes
that $\mathcal{M}\,;\,w\models \varphi$ does not hold):

@@

\paragraph{Syntax and semantics of modal logics}
To implement our library, we use a \emph{deep embedding} of
formulas, i.e. we represent formulas as the following Coq inductive datatype.
\begin{minted}{coq}
Inductive formula: Set :=
| Lit    : nat -> formula
| Neg    : formula -> formula
| Box    : formula -> formula
| Dia    : formula -> formula
| And    : formula -> formula -> formula
| Or     : formula -> formula -> formula
| Implies: formula -> formula -> formula.
\end{minted}
Most constructors of the type have an immediate meaning, since they represent
connectives of propositional logics. The constructor \mintinline{coq}{Lit}
is used to construct propositional variables; constructor \mintinline{coq}{Box}
denotes the modal operator $\Box$ and \mintinline{coq}{Dia}, the operator
$\Diamond$. Furthermore, the type \mintinline{coq}{theory} is defined as a list
of formulas.

@@

Before defining the Kripke semantics, we need to represent the model used to
interpret a modal logic formula. We encode a frame by a record formed by
a type \mintinline{coq}{W: Set} to represent the set of worlds and an
accessibility relation denoted by a type \mintinline{coq}{R: W -> W -> Prop}.
\begin{minted}{coq}
Record Frame: Type := {
  W: Set;
  R: W -> W -> Prop;
}.
\end{minted}
A model is then denoted by another record, formed by a frame and a labeling function,
which assign the truth value of a variable at a given world.
\begin{minted}{coq}
Record Model: Type := {
  F: Frame; 
  v: nat -> (W F) -> Prop; 
}.
\end{minted}
Note that the type of the labeling function \mintinline{coq}{v} ensures that it
refers to a world from the set of possible worlds in the frame
\mintinline{coq}{F}\footnote{We recall that \mintinline{coq}{W F} is Coq's syntax for
accessing field \mintinline{coq}{W} from the record \mintinline{coq}{F}.}. The encoding of the Kripke semantics is done by a Coq function
that implements the relation of Figure~\ref{fig:kripke}. We omit the code
for propositional logics connectives for brevity. The interesting parts are for
variables (constructor \mintinline{coq}{Lit}) which uses the labeling function
\mintinline{coq}{v} from the model \mintinline{coq}{M}. The meaning of
modalities $\Box$ and $\Diamond$ are represented by quantifying over worlds
using the accessibility relation $R$. 
\begin{minted}{coq}
Fixpoint formula_valuation (M: Model) (w: W (F M))
                           (phi: formula): Prop :=
  match phi with
  | Lit     x
     => v M x w (* should be inhabited, i.e., true *)
  | Box     psi
     => forall w': W (F M), R (F M) w w' ->
                            formula_valuation M w' psi
  | Dia     psi
     => exists w': W (F M), R (F M) w w' /\
                            formula_valuation M w' psi
  (* some code omitted *)
  end.
\end{minted}
In order to ease the task of writing formulas and its semantics properties, we
define Coq notations that mimics the standard way of writing them.

\paragraph{The Hilbert deductive system}
Basically, a Hilbert-style deductive system allows the use of four
possibilities in a proof: 1) the use of a hypothesis; 2) the use of an axiom;
3) the use of the modus ponens rule or 4) the use of the generalization (necessity)
rule. The inductive type \mintinline{coq}{deduction} encodes these possibilities
as its constructors. The \mintinline{coq}{deduction} type is parameterized by
the set of axioms that form a system for some modal logic. In this way, our definition can be
used by several systems alike.
\begin{minted}{coq}
Inductive deduction (A: axiom -> Prop):
                          theory -> formula -> Prop :=
  (* Premise. *)
  | Prem: forall (t: theory)
                 (f: formula)
                 (i: nat),
          (nth_error t i = Some f) -> deduction A t f
  (* Axiom. *)
  | Ax: forall (t: theory)
               (a: axiom)
               (f: formula),
        A a -> instantiate a = f -> deduction A t f
  (* Modus Ponens. *)
  | Mp: forall (t: theory)
               (f g: formula)
               (d1: deduction A t (f -> g))
               (d2: deduction A t f),
        deduction A t g
  (* Generalization. *)
  | Nec: forall (t: theory)
                (f: formula)
                (d1: deduction A t f),
         deduction A t (Box f).
\end{minted}
The constructors \mintinline{coq}{Prem}, \mintinline{coq}{Mp} and \mintinline{coq}{Nec}
represent rules for hypothesis, modus ponens and generalization, respectively.
Before discussing the \mintinline{coq}{Ax}, we need to discuss how axioms are
represented in our library. The inductive type \mintinline{coq}{axiom} specifies
the different axioms supported by our library.
\begin{minted}{coq}
Inductive axiom: Set :=
  | ax1  : formula -> formula -> axiom
  | ax2  : formula -> formula -> formula  -> axiom
  (* some code omitted *)
  | axK  : formula -> formula -> axiom
  | axPos: formula -> formula -> axiom
  | axT  : formula -> axiom
  | axB  : formula -> axiom
  | axK4 : formula -> axiom
  | axD  : formula -> axiom
  | axK5 : formula -> axiom.
\end{minted}
Each constructor of type \mintinline{coq}{axiom} represents an axiom and
its type stores the subformulas which compose the axiom. The type has
17 constructors: the first 10 encode axioms for propositional logics
and, the last 7 are for different formulas which are taken as axioms by common
systems of modal logic. Axioms are built using the function \mintinline{coq}{instantiate},
which receives an axiom and builds a corresponding formula by pattern matching as
follows:
\begin{minted}{coq}
Definition instantiate (a: axiom): formula :=
  match a with
  | ax1    p   q
     => [! p -> (q -> p) !]
  | ax2    p   q   r
     => [! (p -> (q -> r)) -> ((p -> q) -> (p -> r)) !]
  | ax3    p   q
     => [! (~ q -> ~ p) -> (p -> q) !]
  | axK    p   q
     => [! [] (p -> q) -> ([] p -> [] q) !]
  | axPos  p   q
     => [! <> (p \/ q) -> (<> p \/ <> q) !]
  (* some code omitted *)
  end.
\end{minted}
Notation \mintinline{coq}{[! p !]} is used to write modal logic formulas using a
syntax closer to mathematical practice using Coq. This is necessary to avoid
parsing ambiguities, since some symbols are also used by Coq's syntax.
Modalities $\Box$ and $\Diamond$ are denoted by \mintinline{coq}{[]}
and \mintinline{coq}{<>}, respectively.

@@

\paragraph{Systems of modal logics} In order to use the deductive system to
check the provability of formulas, we need to specify the axioms that are
included in the system. Our approach is to define predicates
which denote each system. The simplest set of axioms is usually refereed as
\textbf{K} and it includes the axiom $\Box(A \to B) \to (\Box A \to \Box B)$
and all the axioms for propositional logic. We represent system \textbf{K}
by the following inductive type:
\begin{minted}{coq}
Inductive K: axiom -> Prop :=
  | K_ax1: forall p q, K (ax1 p q)
  | K_ax2: forall p q r, K (ax2 p q r)
  (* some code omitted *)
  | K_axK: forall p q, K (axK p q)
\end{minted}
Type \mintinline{coq}{K} is a predicate over axioms: it has one constructor for
each of the valid axioms which form the system. Given a value
of \mintinline{coq}{axiom}, a value of type \mintinline{coq}{K a} denotes a
proof that a given axiom is part of the system \textbf{K}. This is the main
trick used by the constructor \mintinline{coq}{Ax}, of type
\mintinline{coq}{deduction}, to generically represent axioms for
different systems of modal logics uniformly. In order to understand how it
works, let us review its type (below, remember that \mintinline{coq}{A: axiom -> Prop}
denotes a system of axioms for modal logics):
\begin{minted}{coq}
Ax: forall (t: theory)
           (a: axiom)
           (f: formula),
           A a -> instantiate a = f -> deduction A t f
\end{minted}
Note that the first parameter of constructor \mintinline{coq}{Ax} is a value
of type \mintinline{coq}{A a}, which is only inhabited if the predicate
\mintinline{coq}{A} holds for the axiom \mintinline{coq}{a}. Through this
evidence, is possible to create the corresponding axiom formula instance by
using function \mintinline{coq}{instantiate}. In this way, we are able to
represent all systems of axioms for modal logics using a single constructor
for its axioms.

@@

\paragraph{Frame Correspondence} 
As already mentioned, there are some modal formula schemas that correspond
to properties of the accessibility relation of a modal system, if we take them
as an axiom in the system.  While the proof that a certain property in a frame
implies the correspondent axiom is usually  direct, the inverse often needs
more effort. On the Coq proofs developed for this work,  the former case is
mostly unfoldings, simplifications and applications of premises
while the latter, apart from the necessity and burden of a proof by contraposition, 
requires the construction of a witness valuation function. 
Nevertheless, all Coq's proofs are fairly similar to the respective
``pen and paper'' proofs found in the literature. 
Our work proves correspondence for the following properties: reflexivity,
transitivity, symmetry,  euclidean, seriality, functionality, density, convergency,
and converse well-foundness.

@@

\paragraph{Formalizations of modal logics on proof assistants}

The formalization of epistemic modal logic in Coq was the subject of Wind's
work~\cite{dewind}.
The author made a Coq implementation of modal logic based on the S5 system,
which consists of  the axioms $K$, $T$, $4$ and $5$. The objective was the
modeling of two well known puzzles.
Wind's formalization was separated into some components. The first consists in
defining the syntax of formulas and its Kripke semantics. Next, the author uses
their definitions to give some examples of modal operators, worlds and how they
interact with each other. Using the built machinery, the author developed
introduction and elimination rules for modal logic operators and use them to
model the wize man's and muddy children puzzles.

@@

The development of a completeness proof for modal logic S5 in Lean proof
assistant was the subject of the Bentzen's work~\cite{Bentzen2021}. The author
follows the approach due to Henkin for S5 completeness and it depends on
the \texttt{mathlib}, the Lean's library of mathematical results~\cite{mathlib20}.

Maggesi et. al. developed a formalization of the completeness proof for GL using
the HOL light theorem prover~\cite{Maggesi21}. The authors argue that using the
standard Henkin strategy does not work in GL without modifications. The core idea of
Maggesi et. al. strategy was building maximal consistent lists in terms of
components of the given formula non-proven in the calculus.

The following table summarizes the formalized modal logics systems by our library
and related works:
\vspace{0.5cm} {
  \begin{table}[H]
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
      \hline
      \multirow{2}{*}{Authors} & \multicolumn{9}{c|}{Systems considered} \\ \cline{2-10} 
                               & K & D & B & T & K4 & K5 & S4 & S5 & GL\\ \hline
      Wind et. al.~\cite{dewind}     & X &   &   & X & X  & X  & X  &  X & \\ \hline
      Doczkal et. al.~\cite{saarland}   & X &   &   & X & X  &    &    &  &  \\ \hline
      Benzmüller et. al.~\cite{Chris_Bruno}& X &   & X & X &  X &    &    & X &  \\ \hline
      Bentzen~\cite{Bentzen2021} & X & & & X & & X & X & X & \\ \hline
      Maggesi et. al.~\cite{Maggesi21} & X & & & & & & & & X \\ \hline
      This work         & X & X & X & X & X  &  X &  X & X  & X \\ \hline
    \end{tabular}
    \centering
    \caption{Comparision with different Coq modal logics developements.}
    \label{tab:comparacao_trabalhos_rel}
  \end{table}
}

@@

\paragraph{Applications of modal logics in semantics} Another application of
modal logics is its use for reasoning about programming languages semantics.
Dynamic logic was introduced by Pratt in late 70's as modal logic for reasoning
about programs~\cite{Pratt76,Pratt80}. The key idea on dynamic logic is to express the
Hoare triples by a modality meaning ``after a program $p$ executes, formula
$\varphi$ will be true''. The programming languages considered in Pratt's
original work are composed by assignment and regular language operators
(union, concatenation and Kleene star)~\cite{Pratt76}.

An application of modal logic IS5 as a type system for a distributed
programming language was proposed in~\cite{Murphy04}. 
In that work, modalities are used to represent types for mobile code
and remote addresses. Murphy et. al defined the language by interpreting
the possible worlds of the intutionistic modal logic IS5 to be nodes on a
network and the modalities $\Box$ and $\Diamond$ to reflect mobility and
locality, respectively. Authors develop an operational semantics and show
that their formalization is type-safe.

The application of a multimodal separation logic for the semantics of low-level languages
was the subject of Dockins et. al. work~\cite{Dockins08}. In that work, authors
combine several modal operators that do not need to satisfy the same frame
conditions and seperation logics for reasoning about non-trivial subsets of
low-level programming languages. Authors developed their formalization through a
shallow embedding in Coq proof assistant and provide two case studies:
the verification of a low-level virtual machine and a C-like language.

A model for recursive, impredicatively quantified types with mutable references
was studied by Appel et. al.~\cite{Appel07}. Their interest in formalizing such
advanced typing concepts was in the context of typed assembly languages.
The authors stabilished a semantic type soundness proof by defining a Kripke-style 
semantics of the G\"odel-L\"ob logic of provability. All their definitions were
also mechanized through a shallow embedding in Coq.

@@

In this work, we presented a sound deep embedding of arbitrary
normal modal logics in Coq. Our approach is able to represent deduction
uniformly over different modal logic systems by using different axiom sets
as a predicate parameter in the deduction inductive type. Using this generic
representation of deduction we are able to prove the soundness theorem and
apply it on proving L\"ob's theorem by assuming fixed points. % All source code in this article has
% been formalized in Coq version 8.15.0 and it is available at~\cite{modal-coq-lib}.

>>>>

Formally, we can only interpret a formula with respect to a \emph{model}, which is
defined as follows.
A \emph{frame} $\mathcal{F} = \langle W, R \rangle$ consists of a pair formed by
a non-empty set of worlds, $W$, and a binary accessibility relation between
worlds, $R \subseteq W \times W$. A pair $w_1Rw_2$ denotes that world $w_2$ is
accessible from $w_1$. A model $\mathcal{M} = \langle \mathcal{F}, \mathcal{V} \rangle$
is formed by a frame $\mathcal{F}$ and a total labeling function
$\mathcal{V} : \mathcal{P}\to 2^W$, which indicates the set of worlds where each propositional variable holds, i.e. when $w \in \mathcal{V}(A)$
for some $A \in\mathcal{P}$, we say that $A$ holds in $w$. The satisfiability of a formula
in a given world is defined by inductive relation $\mathcal{M}\,;\,w\models \varphi$ as
follows (notation $\mathcal{M}\,;\,w\not\models \varphi$ denotes
that $\mathcal{M}\,;\,w\models \varphi$ does not hold):

@@

\paragraph{Syntax and semantics of modal logics}
To implement our library, we use a \emph{deep embedding} of
formulas, i.e. we represent formulas as the following Coq inductive datatype.
\begin{minted}{coq}
Inductive formula: Set :=
| Lit    : nat -> formula
| Neg    : formula -> formula
| Box    : formula -> formula
| Dia    : formula -> formula
| And    : formula -> formula -> formula
| Or     : formula -> formula -> formula
| Implies: formula -> formula -> formula.
\end{minted}
Most constructors of the type have an immediate meaning, since they represent
connectives of propositional logic. The constructor \mintinline{coq}{Lit}
is used to construct propositional variables; constructor \mintinline{coq}{Box}
denotes the modal operator $\Box$ and \mintinline{coq}{Dia}, the operator
$\Diamond$. Furthermore, the type \mintinline{coq}{theory} is defined as a list
of formulas.

@@

Before defining the Kripke semantics, we need to represent the model used to
interpret a modal logic formula. We encode a frame by a record formed by
a type \mintinline{coq}{W: Set} to represent the set of worlds and an
accessibility relation denoted by a type \mintinline{coq}{R: W -> W -> Prop}.
\begin{minted}{coq}
Record Frame: Type := {
  W: Set;
  R: W -> W -> Prop;
}.
\end{minted}
A model is then denoted by another record, formed by a frame and a labeling function,
which assigns the truth value of a variable at a given world.
\begin{minted}{coq}
Record Model: Type := {
  F: Frame; 
  v: nat -> (W F) -> Prop; 
}.
\end{minted}
Note that the type of the labeling function \mintinline{coq}{v} ensures that it
refers to a world from the set of possible worlds in the frame
\mintinline{coq}{F}\footnote{We recall that \mintinline{coq}{W F} is Coq's syntax for
accessing field \mintinline{coq}{W} from the record \mintinline{coq}{F}.}. The encoding of the Kripke semantics is done by a Coq function
that implements the relation of Figure~\ref{fig:kripke}. We omit the code
for propositional logics connectives for brevity. The interesting parts are for
variables (constructor \mintinline{coq}{Lit}) which uses the labeling function
\mintinline{coq}{v} from the model \mintinline{coq}{M}. The meaning of
modalities $\Box$ and $\Diamond$ is represented by quantifying over worlds
using the accessibility relation $R$. 
\begin{minted}{coq}
Fixpoint formula_valuation (M: Model) (w: W (F M))
                           (phi: formula): Prop :=
  match phi with
  | Lit     x
     => v M x w (* should be inhabited, i.e., true *)
  | Box     psi
     => forall w': W (F M), R (F M) w w' ->
                            formula_valuation M w' psi
  | Dia     psi
     => exists w': W (F M), R (F M) w w' /\
                            formula_valuation M w' psi
  (* some code omitted *)
  end.
\end{minted}
In order to ease the task of writing formulas and its semantics properties, we
define Coq notations that mimic the standard way of writing them.

\paragraph{The Hilbert deductive system}
Basically, a Hilbert-style deductive system allows the use of four
possibilities in a proof: 1) the use of a hypothesis; 2) the use of an axiom;
3) the use of the modus ponens rule or 4) the use of the generalization (necessity)
rule. The inductive type \mintinline{coq}{deduction} encodes these possibilities
as its constructors. The \mintinline{coq}{deduction} type is parameterized by
the set of axioms that form a system for some modal logic. In this way, our definition can be
used by several systems alike.
\begin{minted}{coq}
Inductive deduction (A: axiom -> Prop):
                          theory -> formula -> Prop :=
  (* Premise. *)
  | Prem: forall (t: theory)
                 (f: formula)
                 (i: nat),
          (nth_error t i = Some f) -> deduction A t f
  (* Axiom. *)
  | Ax: forall (t: theory)
               (a: axiom)
               (f: formula),
        A a -> instantiate a = f -> deduction A t f
  (* Modus Ponens. *)
  | Mp: forall (t: theory)
               (f g: formula)
               (d1: deduction A t (Implies f g))
               (d2: deduction A t f),
        deduction A t g
  (* Generalization. *)
  | Nec: forall (t: theory)
                (f: formula)
                (d1: deduction A t f),
         deduction A t (Box f).
\end{minted}
The constructors \mintinline{coq}{Prem}, \mintinline{coq}{Mp} and \mintinline{coq}{Nec}
represent rules for hypothesis, modus ponens and generalization, respectively.
Before discussing the \mintinline{coq}{Ax} constructor, we need to discuss how axioms are
represented in our library. The inductive type \mintinline{coq}{axiom} specifies
the different axioms supported by our library.
\begin{minted}{coq}
Inductive axiom: Set :=
  | ax1  : formula -> formula -> axiom
  | ax2  : formula -> formula -> formula  -> axiom
  (* some code omitted *)
  | axK  : formula -> formula -> axiom
  | axPos: formula -> formula -> axiom
  | axT  : formula -> axiom
  | axB  : formula -> axiom
  | axK4 : formula -> axiom
  | axD  : formula -> axiom
  | axK5 : formula -> axiom.
\end{minted}
Each constructor of type \mintinline{coq}{axiom} represents an axiom scheme and
its type stores the subformulas which compose the axiom. The type has
17 constructors: the first 10 encode axioms for propositional logic
and, the last 7 are for different formulas which are taken as axioms by common
systems of modal logic. Axioms are built using the function \mintinline{coq}{instantiate},
which receives an axiom and builds a corresponding formula by pattern matching as
follows:
\begin{minted}{coq}
Definition instantiate (a: axiom): formula :=
  match a with
  | ax1    p   q
     => [! p -> (q -> p) !]
  | ax2    p   q   r
     => [! (p -> (q -> r)) -> ((p -> q) -> (p -> r)) !]
  | ax3    p   q
     => [! (~ q -> ~ p) -> (p -> q) !]
  | axK    p   q
     => [! [] (p -> q) -> ([] p -> [] q) !]
  | axPos  p   q
     => [! <> (p \/ q) -> (<> p \/ <> q) !]
  (* some code omitted *)
  end.
\end{minted}
Notation \mintinline{coq}{[! p !]} is used to write modal logic formulas using a
syntax closer to mathematical practice using Coq. This is necessary to avoid
parsing ambiguities, since some symbols are also used by Coq's syntax.
Modalities $\Box$ and $\Diamond$ are denoted by \mintinline{coq}{[]}
and \mintinline{coq}{<>}, respectively.

@@

\paragraph{Systems of modal logics} In order to use the deductive system to
check the provability of formulas, we need to specify the axioms that are
included in the system. Our approach is to define predicates
which denote each system. The simplest set of axioms is usually referred as
\textbf{K} and it includes the axiom $\Box(A \to B) \to (\Box A \to \Box B)$
and all the axioms for propositional logic. We represent system \textbf{K}
by the following inductive predicate:
\begin{minted}{coq}
Inductive K: axiom -> Prop :=
  | K_ax1: forall p q, K (ax1 p q)
  | K_ax2: forall p q r, K (ax2 p q r)
  (* some code omitted *)
  | K_axK: forall p q, K (axK p q)
\end{minted}
Type \mintinline{coq}{K} is a predicate over axioms: it has one constructor for
each of the valid axioms which form the system. Given a value
of \mintinline{coq}{axiom}, a value of type \mintinline{coq}{K a} denotes a
proof that a given axiom is part of the system \textbf{K}. This is the main
trick used by the constructor \mintinline{coq}{Ax}, of type
\mintinline{coq}{deduction}, to generically represent axioms for
different systems of modal logics uniformly. In order to understand how it
works, let us review its type (below, remember that \mintinline{coq}{A: axiom -> Prop}
denotes a system of axioms for modal logics):
\begin{minted}{coq}
Ax: forall (t: theory)
           (a: axiom)
           (f: formula),
           A a -> instantiate a = f -> deduction A t f
\end{minted}
Note that the fourth parameter of constructor \mintinline{coq}{Ax} is a value
of type \mintinline{coq}{A a}, which is only inhabited if the predicate
\mintinline{coq}{A} holds for the axiom \mintinline{coq}{a}. Through this
evidence, is possible to create the corresponding axiom formula instance by
using function \mintinline{coq}{instantiate}. In this way, we are able to
represent all systems of axioms for modal logics using a single type
for its axioms.

@@

\paragraph{Frame Correspondence} 
As already mentioned, there are some modal formula schemas that correspond
to properties of the accessibility relation of a modal system, if we take them
as an axiom in the system.  While the proof that a certain property in a frame
implies the correspondent axiom is usually  direct, the inverse often needs
more effort. In the Coq proofs developed for this work,  the former case is
mostly unfoldings, simplifications and applications of premises
while the latter, apart from the necessity and burden of a proof by contraposition, 
requires the construction of a witness valuation function. 
Nevertheless, all Coq proofs are fairly similar to the respective
``pen and paper'' proofs found in the literature. 
Our work proves correspondence for the following properties: reflexivity,
transitivity, symmetry,  euclidean, seriality, functionality, density, convergency,
and converse well-foundness.

@@

\paragraph{Formalizations of modal logics in proof assistants}

The formalization of epistemic modal logic in Coq was the subject of de Wind's
work~\cite{dewind}.
The author made a Coq implementation of modal logic based on the S5 system,
which consists of  the axioms $K$, $T$, $4$ and $5$. The objective was the
modeling of two well known puzzles.
De Wind's formalization was separated into some components. The first consists in
defining the syntax of formulas and its Kripke semantics. Next, the author uses
their definitions to give some examples of modal operators, worlds and how they
interact with each other. Using the built machinery, the author developed
introduction and elimination rules for modal logic operators and use them to
model the wize man's and muddy children puzzles.

@@

The development of a completeness proof for modal logic S5 in Lean proof
assistant was the subject of the Bentzen's work~\cite{Bentzen2021}. The author
follows the approach due to Henkin for S5 completeness and it depends on
\texttt{mathlib}, the Lean's library of mathematical results~\cite{mathlib20}.

Maggesi et al. developed a formalization of the completeness proof for GL using
the HOL light theorem prover~\cite{Maggesi21}. The authors argue that using the
standard Henkin strategy does not work in GL without modifications. The core idea of
Maggesi et al. strategy was building maximal consistent lists in terms of
components of the given formula non-proven in the calculus.

The following table summarizes the formalized modal logics systems by our library
and related works:
\vspace{0.5cm} {
  \begin{table}[H]
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
      \hline
      \multirow{2}{*}{Authors} & \multicolumn{9}{c|}{Systems considered} \\ \cline{2-10} 
                               & K & D & B & T & K4 & K5 & S4 & S5 & GL\\ \hline
      Wind et al.~\cite{dewind}     & X &   &   & X & X  & X  & X  &  X & \\ \hline
      Doczkal et al.~\cite{saarland}   & X &   &   & X & X  &    &    &  &  \\ \hline
      Benzmüller et al.~\cite{Chris_Bruno}& X &   & X & X &  X &    &    & X &  \\ \hline
      Bentzen~\cite{Bentzen2021} & X & & & X & & X & X & X & \\ \hline
      Maggesi et al.~\cite{Maggesi21} & X & & & & & & & & X \\ \hline
      This work         & X & X & X & X & X  &  X &  X & X  & X \\ \hline
    \end{tabular}
    \centering
    \caption{Comparision with different Coq modal logics developements.}
    \label{tab:comparacao_trabalhos_rel}
  \end{table}
}

@@

\paragraph{Applications of modal logics in semantics} Another application of
modal logics is its use for reasoning about programming languages semantics.
Dynamic logic was introduced by Pratt in late 70's as modal logic for reasoning
about programs~\cite{Pratt76,Pratt80}. The key idea of dynamic logic is to express the
Hoare triples by a modality meaning ``after a program $p$ executes, formula
$\varphi$ will be true''. The programming languages considered in Pratt's
original work are composed by assignment and regular language operators
(union, concatenation and Kleene star)~\cite{Pratt76}.

An application of modal logic IS5 as a type system for a distributed
programming language was proposed in~\cite{Murphy04}. 
In that work, modalities are used to represent types for mobile code
and remote addresses. Murphy et al defined the language by interpreting
the possible worlds of the intutionistic modal logic IS5 to be nodes on a
network and the modalities $\Box$ and $\Diamond$ to reflect mobility and
locality, respectively. Authors develop an operational semantics and show
that their formalization is type-safe.

The application of a multimodal separation logic for the semantics of low-level languages
was the subject of Dockins et al.'s work~\cite{Dockins08}. In that work, authors
combine several modal operators that do not need to satisfy the same frame
conditions and seperation logics for reasoning about non-trivial subsets of
low-level programming languages. Authors developed their formalization through a
shallow embedding in Coq proof assistant and provide two case studies:
the verification of a low-level virtual machine and a C-like language.

A model for recursive, impredicatively quantified types with mutable references
was studied by Appel et al.~\cite{Appel07}. Their interest in formalizing such
advanced typing concepts was in the context of typed assembly languages.
The authors stabilished a semantic type soundness proof by defining a Kripke-style 
semantics of the G\"odel-L\"ob logic of provability. All their definitions were
also mechanized through a shallow embedding in Coq.

@@

In this work, we presented a sound deep embedding of arbitrary
normal modal logics in Coq. Our approach is able to represent deduction
uniformly over different modal logic systems by using different axiom sets
as a predicate parameter in the deduction inductive type. Using this generic
representation of deduction we are able to prove a soundness theorem, and, in
order to demonstrate the applicability of the proposed work, prove L\"ob's theorem
by assuming fixed points. % All source code in this article has
% been formalized in Coq version 8.15.0 and it is available at~\cite{modal-coq-lib}.
